---
format: html
editor: visual
author: Mahsa Mohammadi, Luka Milosevic
---

```{r, eval = F}
# This code chunk contains code to install all of the dependencies
# necessary to compile and run your report, using if-statements to
# reduce install time for unnecessary code.
# It should be set to eval = F by default, so you aren't installing
# software on someone's computer without their consent.

# This works for packages that are on CRAN
if (!"tidyverse" %in% installed.packages()) {
  install.packages("tidyverse")
}
if (!"remotes" %in% installed.packages()) {
  install.packages("remotes")
}

# This is how to do it for a package that's only on github
if (!"emo" %in% installed.packages()) {
  remotes::install_github("hadley/emo")
install.packages("plotly")
}
if (!requireNamespace("igraph", quietly = TRUE)) {
  install.packages("igraph")
}
library(igraph)

if (!requireNamespace("viridisLite", quietly = TRUE)) {
  install.packages("viridisLite")
}
library(viridisLite)
if (!requireNamespace("viridis", quietly = TRUE)) {
  install.packages("viridis")
}
library(viridis)

if (!requireNamespace("skimr", quietly = TRUE)) {
  install.packages("skimr")
}
library(skimr)

```

# Introduction

The dataset, sourced from [Hollywood age gap site](https://hollywoodagegap.com/), provides information on movies and the age gaps between the love interests portrayed by actors. The variables include movie name, movie release year, director, age difference, number of couples, names and of the actors involved in romance, their birth dates and age, as well as the gender of character they are portraying in the movie. Given its primary focus on significant age differences between actors portraying characters in love, the dataset does not exhibit anomalies. It is designed to highlight and analyze substantial variations in age within the context of romantic relationships depicted in films.

The unique aspect of this dataset is its focus on age gaps in romantic relationships within movies. The dataset is curated based on specific rules, ensuring that the actors portrayed actual love interests, with the youngest being at least 17 years old. Additionally, the dataset excludes animated characters, emphasizing real-life portrayals of romantic relationships in films.

# Methods

We utilized the tidyverse package, specifically relaying on the functionalities of the `dplyr` and `lubridate` packages for effective data manipulation. Tidyverse is a comprehensive suite designed to streamline data analysis and visualization workflows. Within this framework, dplyr provides versatile tools for data manipulation tasks.

Our dataset was already organized by the principles of tidy data, which advocates for a standardized organization of datasets. This organizational structure involves arranging variables in columns, observations in rows, and maintaining a singular observational unit in each table. The structured format of our data minimized the need for extensive manipulation.

The `dplyr` package's `mutate` function played a central role in our data manipulation process. This function facilitated the creation and modification of variables within the dataset, enabling the seamless introduction of calculated columns tailored to our analytical needs. Combined with `ymd` function from `lubridate` package, we have successfully ensured the transformation of the birth date variable into a standardized Date type within the dataset. Additionaly, we emplyed `skim` fucntion from `skimr` package to derive a simple overveiw of the data.

```{r, results = 'hide'}
#| echo: false
dat = read.csv("age_gaps.csv")
library(tidyverse)
library(skimr)

glimpse(dat)

dat %>%
  mutate(actor_1_birthdate = ymd(actor_1_birthdate),
         actor_2_birthdate = ymd(actor_2_birthdate), 
         decade = 10 * floor(release_year / 10)) -> dat

dat_long <- dat %>%
  gather(key = "actor", value = "age", actor_1_age, actor_2_age)
```

# Topic of Exploration

##  Descriptive Statistics:

-   Calculate basic statistics such as mean, median, and standard deviation for age and age gap.
    
```{r, results = 'hide'}
    skimsummary = skim (dat)
    summary(dat)
```    

The dataset contains 1,155 entries, with movies spanning from 1935 to 2022. The average age difference between actors is 10.42 years, with a minimum of 0 years and a maximum of 52 years. The dataset features up to 7 couples per movie, with a mean of 1.398 couples. The actors' birthdates indicate a range from 1889 to 1996, and the average age of actor 1 is 40.64 years, while actor 2 has an average age of 30.21 years. The dataset covers movies released in different decades, with a mean release year of 2001. Overall, these statistics provide a comprehensive overview of the age dynamics, release timelines, and other characteristics within the movie dataset.

```{r}
age <- c(dat$actor_1_age, dat$actor_2_age)
age <- as.data.frame(age)

# Create a histogram 
ggplot(age, aes(x=age)) +
  geom_histogram(binwidth=1, position='identity',fill = "skyblue", alpha=0.5, color='black') +
  labs(title='Histogram of Actor Ages', x='Age', y='Frequency') +
  theme_minimal()
```

The average of the age is 35 years old, while the youngest actor is 17 and oldest one is 81 years old.

-   Create histograms or box plots to visualize the distribution of ages and age gaps.

```{r}
library(plotly)

gg <- ggplot(dat, aes(x = age_difference)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Age Gap Distribution",
       x = "Age Difference",
       y = "Frequency") +
  scale_x_continuous(breaks = seq(0, 55, by = 5)) +
  theme_minimal()

plotly_plot <- ggplotly(gg)

plotly_plot
```

Based on this graph, the age gap between actors is mostly about 2-3 years. But there are some out liars that age gap is more than 40 years!

##  Relationship Dynamics:

-   Explore the correlation between age and age gap. Are there trends or patterns that suggest certain age differences are more common?

```{r}
    ggplot(dat, aes(x = actor_1_age, y = age_difference)) +
      geom_point(alpha = 0.7, color = 'skyblue') +
      labs(title = 'Relationship Between Age and Age Gap',
           x = 'Actor 1 age',
           y = 'Age Gap') +
      theme_minimal()

    ggplot(dat, aes(x = actor_2_age, y = age_difference)) +
      geom_point(alpha = 0.7, color = 'skyblue') +
      labs(title = 'Relationship Between Age and Age Gap',
           x = 'Actor 2 age',
           y = 'Age Gap') +
      theme_minimal()
```

This graphs are based on actor 1 age and age gap and actor 2 age and age gap. These 2 figures are kind of mirrored each other, since actor 1 in our dataset is always the older partner.

##  Temporal Analysis:

-   Examine if there are trends over time. For example, has there been a change in the average age gap between actors over the years?

In this case, we will group the movies by release decade.

```{r, warning=FALSE}
ggplot(dat, aes(x = decade, y = age_difference)) +
  geom_point(alpha = 0.7, color = "skyblue", size = 3) +
  geom_smooth(method = "loess", se = TRUE, color = "blue", linetype = "solid", size = 1) +
  labs(title = "Scatter Plot of Decade vs. Age Difference with Smoothed Line",
       x = "Decade",
       y = "Age Difference") +
  scale_x_continuous(breaks = seq(1930, 2020, by = 10)) +
  theme_minimal() +
  theme(panel.grid = element_line(color = "lightgray", linetype = "dashed"))
```

This graph shows the highest age gaps are in 1990 and 2000. the overall age gap trend is more or less constant through the decades.


##  Network Analysis:

-   Construct a network graph representing relationships between actors. Analyze the network structure, identifying key actors who are connected to many others.

```{r}
# Filter the DataFrame for a specific director
library(igraph)
director_name <- "Woody Allen"
filtered_data <- dat[dat$director == director_name, ]

# Create a network graph
graph <- graph.data.frame(filtered_data[, c("actor_1_name", "actor_2_name")], directed = FALSE)

# Set up plot parameters
par(mfrow=c(1,1), mar = c(0,0,0,0))
layout <- layout_with_kk(graph)
color_gradient <- viridisLite::viridis(100, option = "plasma")
color_mapping <- color_gradient[cut(filtered_data$age_difference, breaks = 100, labels = FALSE)]

# Plot the network graph with adjusted parameters, layout, and color mapping
plot(graph, 
     main = paste("Actor Network for Director:", director_name), 
     edge.arrow.size = 0.5,
     vertex.size = 20,
     vertex.label.cex = 0.8,
     layout = layout,
     edge.color = color_mapping
)
# Add legend
legend("bottomright", 
       legend = levels(cut(filtered_data$age_difference, breaks = 100)), 
       fill = color_gradient,
       title = "Age Difference",
       cex = 0.7)

```

I didn't know that Woody Allen played in his movie too! Also it looks like he was in the relationship.

```{r}
# Filter the DataFrame for a specific director
director_name <- "Guy Hamilton"
filtered_data <- dat[dat$director == director_name, ]

# Create a network graph
graph <- graph.data.frame(filtered_data[, c("actor_1_name", "actor_2_name")], directed = FALSE)

par(mfrow=c(1,1), mar = c(0,0,0,0))

#Choose a layout algorithm

layout <- layout_with_kk(graph)
# Create a color gradient based on age difference
color_gradient <- viridisLite::viridis(100, option = "plasma")

# Map age difference to colors
color_mapping <- color_gradient[cut(filtered_data$age_difference, breaks = 100, labels = FALSE)]

# Plot the network graph with adjusted parameters, layout, and color mapping
plot(graph, 
     main = paste("Actor Network for Director:", director_name), 
     edge.arrow.size = 0.5,
     vertex.size = 20,
     vertex.label.cex = 0.8,
     layout = layout,
     edge.color = color_mapping
)
# Add legend
legend("bottomright", 
       legend = levels(cut(filtered_data$age_difference, breaks = 100)), 
       fill = color_gradient,
       title = "Age Difference",
       cex = 0.7)
```

##  Gender Analysis:

-   Analyze if there are gender-based differences in age gaps.

```{r}
    gender_age_diff = dat %>% 
       group_by(character_1_gender) %>%
       summarise(mean_age_diff_gender = mean(age_difference))
    man_age_diff = gender_age_diff %>% 
      filter(character_1_gender == "man") 
    woman_age_diff = gender_age_diff %>% 
      filter(character_1_gender =="woman")

    t_test_result = t.test(age_difference ~ character_1_gender, data = dat)
    p.value=t_test_result$p.value
```

The analysis indicates a significant difference in age gaps between male and female characters. For example, the average age gap when the male character is older is `r round(man_age_diff$mean_age_diff_gender, 1)` years, while in cases when the female character is older, the average age difference is `r round(woman_age_diff$mean_age_diff_gender, 1)` years. After performing the t-test, we can conclude that this difference is highlight significant (p\>0.0001).

##  Regression Analysis:

-   Build regression models to predict age gap based on various factors such as the age of older actor and movie release year.

```{r, results='hide'}
model <- lm(age_difference ~ actor_1_age + decade, data = dat)
summary(model)
```

```{r}
ggplot(dat, aes(x = actor_1_age, y = age_difference)) +
  geom_point(color="skyblue") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title = "Prediction of the Age difference based on Older Actor Age through decades",
       x = "Older actor Age",
       y = "Age Difference") +
  facet_wrap(~decade) +
  theme_minimal()
```


Then, you can reference @fig-iris in the text and the appropriate cross-reference will be generated.

You can find additional information about formatting figures generated from code in the [quarto documentation](https://quarto.org/docs/authoring/figures.html#computations).

# Additional Exploration topic

1.   Which directors have the highest average age differences in their movies?

```{r}
dat %>%
  group_by(director) %>%
  summarize(mean_age_diff = mean(age_difference)) %>%
  arrange(desc(mean_age_diff)) %>% 
  head(5) %>% 
  pull(director) -> top_directors
```

Five directors with the highest age difference couples in their movies are `r top_directors`. However, recognizing that many directors have only one couple in their movies, we can refine our analysis by sub-setting the data to include only directors with at least three couples. Afterward, we can calculate the average age difference for this subset of directors.

```{r}
dat %>%
  filter(couple_number>=3) %>% 
  group_by(director) %>%
  summarize(mean_age_diff = mean(age_difference)) %>%
  arrange(desc(mean_age_diff)) %>% 
  head (5) %>% 
  pull(director) -> top_directors3
```

Top 5 directors with the highest age difference and at least 3 couples are `r top_directors3`.


2.  We might be curious to see what are the movies? Do they have specific theme?

```{r}
dat %>%
  filter(couple_number>=3) %>% 
  group_by(movie_name) %>%
  summarize(mean_age_diff = mean(age_difference)) %>%
  arrange(desc(mean_age_diff)) %>% 
  head (5) %>% 
  pull(movie_name) -> movie_name
```

Based on the fact that 3 out of 5 movies with the most couples (`r movie_name`) are about the most famous MI6 agent with the licence to kill, we can conclude that James Bond is the Casanova of 20^th^ century.

3.  Who is the most active director?

```{r}
directors <- table(dat$director)

top_10 <- head(sort(directors, decreasing = TRUE), 10)
top_10=as.data.frame(top_10)
```

The most active directors are `r top_10$Var1`.

# Conclusion

Here, you want to summarize the main points of what you've learned from this investigation, in paragraph form.

## Tips

(delete this section from your report!)

Almost anything you might want to know about how to format output in quarto can be found [here](https://quarto.org/docs/authoring/markdown-basics.html). Feel free to email/come to office hours to figure out how to do XYZ - part of the goal of making you write this report is that I want you to know how to write e.g. a journal paper in Quarto as well, so now's the time to experiment.

If you want to know what the wordcount of your report is, you can run the following command in your terminal:

```         
pandoc --lua-filter wordcount.lua report.qmd
```

Notice that I have not pushed `_output/report.html` or the `_output/report_files/` folder to github - this is intentional. I have actually set `_output` to not show up in git, to encourage you all to NOT push the rendered files to github and to instead work from the markdown files directly.

You may find it cleaner to create a figure subdirectory and store any figures that aren't created by R/Python in that folder. I encourage you to organize this repository in a sensible way.
