---
format: html
editor: visual
title: Marriage story
authors: Mahsa Mohammadi, Luka Milosevic
---

```{r, eval = F}
# This code chunk contains code to install all of the dependencies
# necessary to compile and run your report, using if-statements to
# reduce install time for unnecessary code.
# It should be set to eval = F by default, so you aren't installing
# software on someone's computer without their consent.

# This works for packages that are on CRAN
if (!"tidyverse" %in% installed.packages()) {
  install.packages("tidyverse")
}
if (!"remotes" %in% installed.packages()) {
  install.packages("remotes")
}

# This is how to do it for a package that's only on github
if (!"emo" %in% installed.packages()) {
  remotes::install_github("hadley/emo")
install.packages("plotly")
}
if (!requireNamespace("igraph", quietly = TRUE)) {
  install.packages("igraph")
}
library(igraph)

if (!requireNamespace("viridisLite", quietly = TRUE)) {
  install.packages("viridisLite")
}
library(viridisLite)
if (!requireNamespace("viridis", quietly = TRUE)) {
  install.packages("viridis")
}
library(viridis)

if (!requireNamespace("skimr", quietly = TRUE)) {
  install.packages("skimr")
}
library(skimr)
```

# Introduction

The dataset, sourced from [Hollywood age gap site](https://hollywoodagegap.com/), provides information on movies and the age gaps between the love interests portrayed by actors. The variables include movie name, movie release year, director, age difference, number of couples, names and of the actors involved in romance, their birth dates and age, as well as the gender of character they are portraying in the movie. Given its primary focus on significant age differences between actors portraying characters in love, the dataset does not exhibit anomalies. It is designed to highlight and analyze substantial variations in age within the context of romantic relationships depicted in films.

The unique aspect of this dataset is its focus on age gaps in romantic relationships within movies. The dataset is curated based on specific rules, ensuring that the actors portrayed actual love interests, with the youngest being at least 17 years old. Additionally, it excludes animated characters, emphasizing real-life portrayals of romantic relationships in films.

# Methods

We utilized the tidyverse package, specifically relaying on the functionality of the `dplyr` and `lubridate` packages for effective data manipulation. Tidyverse is a comprehensive suite designed to streamline data analysis and visualization workflows. Within this framework, dplyr provides versatile tools for data manipulation tasks.

Our dataset was already organized by the principles of tidy data, which advocates for a standardized organization of datasets. This organizational structure involves arranging variables in columns, observations in rows, and maintaining a singular observational unit in each table. The structured format of our data minimized the need for extensive manipulation.

The `dplyr` package's `mutate` function played a central role in our data manipulation process. This function facilitated the creation and modification of variables within the dataset, enabling the seamless introduction of calculated columns tailored to our analytical needs. Combined with `ymd` function from `lubridate` package, we have successfully ensured the transformation of the birth date variable into a standardized Date type within the dataset. Additionally, we employed `skim` function from `skimr` package to derive a simple overview of the data.

In the gender gap analysis, age differences within the dataset were explored, specifically examining scenarios where the older actor was male compared to female. T-test was utilized to systematically assess whether significant distinctions existed in age based on the gender of the older actor. This analysis provided insights into potential age-related disparities within the dataset.

Additionally, age differences were investigated across decades of movie release year using the `loess` regression model. This approach facilitated the identification of trends and patterns in age discrepancies over time. The use of loess regression allowed for the visualization of potential nonlinearities or trends not immediately apparent through simpler analyses.

To further understand the relationship between the age of older actors and age differences across decades, a linear model was utilized. This analytical tool quantified and described the nature of this relationship, providing a comprehensive perspective on how age differences varied with the older actor's age within each decade.

```{r, results = 'hide'}
#| echo: false
dat = read.csv("age_gaps.csv")
library(tidyverse)
library(skimr)

glimpse(dat)

dat %>%
  mutate(actor_1_birthdate = ymd(actor_1_birthdate),
         actor_2_birthdate = ymd(actor_2_birthdate), 
         decade = 10 * floor(release_year / 10)) -> dat

dat_long <- dat %>%
  gather(key = "actor", value = "age", actor_1_age, actor_2_age)
```

# Topics of Exploration

## Descriptive Statistics:

-   Basic statistical information for age and age gap:

```{r, results = 'hide'}
    skimsummary = skim (dat)
    summary(dat)
```

The dataset contains 1,155 entries, with movies spanning from 1935 to 2022. The average age difference between actors is 10.42 years, with a minimum of 0 years and a maximum of 52 years. The dataset features up to 7 couples per movie, with a mean of 1.398 couples. The actors' birthdates indicate a range from 1889 to 1996, and the average age of actor 1 is 40.64 years, while actor 2 has an average age of 30.21 years. The dataset covers movies released in different decades, with a mean release year of 2001. Overall, these statistics provide a comprehensive overview of the age dynamics, release timelines, and other characteristics within the movie dataset.

```{r}
age <- c(dat$actor_1_age, dat$actor_2_age)
age <- as.data.frame(age)

ggplot(age, aes(x=age)) +
  geom_histogram(binwidth=1, position='identity', fill='skyblue', alpha=0.5, color='black') +
  labs(title='Histogram of actors ages', x='Age', y='Frequency') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 
```

The average of the age is 35 years old, while the youngest actor is 17 and oldest one is 81 years old.

-   Histogram of the distribution of ages and age gaps:

```{r}
library(plotly)

gg <- ggplot(dat, aes(x = age_difference)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Age gap distribution",
       x = "Age difference",
       y = "Frequency") +
  scale_x_continuous(breaks = seq(0, 55, by = 5)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

plotly_plot <- ggplotly(gg)

plotly_plot
```

Based on this graph, the age gap between actors is mostly about 2-3 years. But there are some outliars that age gap is more than 40 years!

## Relationship Dynamics:

-   Is there any correlation between age and age gap?
-   Are there trends or patterns that suggest certain age differences are more common?

```{r}
    ggplot(dat, aes(x = actor_1_age, y = age_difference)) +
      geom_point(alpha = 0.7, color = 'skyblue') +
      labs(title = 'Relationship between age and age gap',
           x = 'Actor 1 age',
           y = 'Age Gap') +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5)) 

    ggplot(dat, aes(x = actor_2_age, y = age_difference)) +
      geom_point(alpha = 0.7, color = 'skyblue') +
      labs(title = 'Relationship between age and age gap',
           x = 'Actor 2 age',
           y = 'Age Gap') +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5)) 
```

The first graph is based on actor 1 age and age gap and the second graph is based on actor 2 age and age gap. These 2 figures are kind of mirrored each other, since actor 1 in our dataset is always the older partner. The graph also validate the point that age gap between actors is mostly under 10 years.

## Temporal Analysis:

-   Is there any trend or change in the average age gap between actors over the years?

To address this question, we will group the movies by their released decade:

```{r, warning=FALSE}
ggplot(dat, aes(x = decade, y = age_difference)) +
  geom_point(alpha = 0.7, color = "skyblue", size = 3) +
  geom_smooth(method = "loess", se = TRUE, color = "blue", linetype = "solid", size = 1) +
  labs(title = "The change in age gap over decades",
       x = "Decade",
       y = "Age Difference") +
  scale_x_continuous(breaks = seq(1930, 2020, by = 10)) +
  theme_minimal() +
  theme(panel.grid = element_line(color = "lightgray", linetype = "dashed"))
```

According to the graph, the highest age gap is present in the movies in 1940s, after which the trend is slowly descending. However, the overall age gap trend is more or less constant through the decades, with slight but probably insignficant differences. Furthermore, the number of movies from the 1930s and 1940s is less than other decades which make it hard to make a general conclusion.

## Network Analysis:

-   Construct a network graph representing the connection between actors and their age gap for specific director or movie.

```{r}
# Filter the DataFrame for a specific director
library(igraph)
director_name <- "Woody Allen"
filtered_data <- dat[dat$director == director_name, ]

# Create a network graph
graph <- graph.data.frame(filtered_data[, c("actor_1_name", "actor_2_name")], directed = FALSE)

# Set up plot parameters
par(mfrow=c(1,1), mar = c(8, 4, 4, 2) + 0.1)  # Adjust the bottom margin (5) as needed
layout <- layout_with_kk(graph)
color_gradient <- viridisLite::viridis(40, option = "plasma")
color_mapping <- color_gradient[cut(filtered_data$age_difference, breaks = 40, labels = FALSE)]

plot(graph, 
     main = paste("Actor Network for Director:", director_name), 
     edge.arrow.size = 0.5,
     vertex.size = 20,
     vertex.label.cex = 0.8,
     layout = layout,
     edge.color = color_mapping
)

# Calculate the legend width based on the graph width
legend_width <- par()$usr[2] - par()$usr[1]  # Graph width

num_colors <- length(color_gradient)
num_columns <- 5  # Adjust the number of columns as needed

legend("bottom", 
       legend = levels(cut(filtered_data$age_difference, breaks = 40)), 
       fill = color_gradient,
       title = "Age Difference",
       cex = 0.7,
       ncol = num_columns,  # Set the number of columns
       xpd = TRUE,  # Allows plotting outside the plot region
       xjust = 0.5,  # Center the legend
       y.intersp = 1.5,  # Adjust the vertical spacing between legend entries
       x = 0,  # Set the x position to place it below the plot
       y = -1.5,  # Set the y position to move it below the plot
       bty = "n")  # Remove the box around the legend

# Reset the xpd parameter to its default value
par(xpd = FALSE)

```

I didn't know that Woody Allen played in his movie too! Also it looks like he was most of the time in the relationship. Rather than him, Most of the other actors in his movies were only in one relationship.

-   Let's look at the network of relationships for another director!

```{r}
# Filter the DataFrame for a specific director
director_name <- "Guy Hamilton"
filtered_data <- dat[dat$director == director_name, ]

# Create a network graph
graph <- graph.data.frame(filtered_data[, c("actor_1_name", "actor_2_name")], directed = FALSE)

par(mfrow=c(1,1), mar = c(8, 4, 4, 2) + 0.1)  # Adjust the bottom margin (5) as needed

#Choose a layout algorithm

layout <- layout_with_kk(graph)
# Create a color gradient based on age difference
color_gradient <- viridisLite::viridis(40, option = "plasma")

# Map age difference to colors
color_mapping <- color_gradient[cut(filtered_data$age_difference, breaks = 40, labels = FALSE)]

# Plot the network graph with adjusted parameters, layout, and color mapping
plot(graph, 
     main = paste("Actor Network for Director:", director_name), 
     edge.arrow.size = 0.5,
     vertex.size = 20,
     vertex.label.cex = 0.8,
     layout = layout,
     edge.color = color_mapping
)
# Add legend
num_colors <- length(color_gradient)
num_columns <- 5  # Adjust the number of columns as needed

# Add legend with multiple columns
legend("bottom", 
       legend = levels(cut(filtered_data$age_difference, breaks = 40)), 
       fill = color_gradient,
       title = "Age Difference",
       cex = 0.7,
       ncol = num_columns,  # Set the number of columns
       xpd = TRUE,  # Allows plotting outside the plot region
       xjust = 0.5,  # Center the legend
       y.intersp = 1.5,  # Adjust the vertical spacing between legend entries
       x = 0,  # Set the x position to place it below the plot
       y = -1.5,  # Set the y position to move it below the plot
       bty = "n")  # Remove the box around the legend

# Reset the xpd parameter to its default value
par(xpd = FALSE)
```

Guy Hamilton's network of the relationships between his actors looks different. They actors created 2 clusters and Roger Moore and Sean Connery are the two center of those clusters with degree of centrality 5 and 3.

## Gender Analysis:

-   Is there any gender-based differences in age gaps?

```{r}
    gender_age_diff = dat %>% 
       group_by(character_1_gender) %>%
       summarise(mean_age_diff_gender = mean(age_difference))
    man_age_diff = gender_age_diff %>% 
      filter(character_1_gender == "man") 
    woman_age_diff = gender_age_diff %>% 
      filter(character_1_gender =="woman")

    t_test_result = t.test(age_difference ~ character_1_gender, data = dat)
    p.value=t_test_result$p.value
```

The analysis indicates a significant difference in age gaps between male and female characters. For example, the average age gap when the male character is older is `r round(man_age_diff$mean_age_diff_gender, 1)` years, while in cases when the female character is older, the average age difference is `r round(woman_age_diff$mean_age_diff_gender, 1)` years. After performing the t-test, we can conclude that this difference is highlight significant (p\>0.0001).

## Regression Analysis:

-   Let's do regression analysis to see if we can find effective variables on predicting the age gap?

```{r, results='hide'}
model <- lm(age_difference ~ actor_1_age + decade, data = dat)
summary(model)
```

```{r}
reg.plot = ggplot(dat, aes(x = actor_1_age, y = age_difference)) +
  geom_point(color="skyblue") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title = "Prediction of the Age difference based on Older Actor Age through decades",
       x = "Older actor Age",
       y = "Age Difference") +
  facet_wrap(~decade) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
print(reg.plot)



models <- lm(age_difference ~ actor_1_age, data = dat)


results_tibble <- tibble()

# Iterate over each unique decade
for (d in unique(dat$decade)) {
  # Fit a linear model for the current decade
  model <- lm(age_difference ~ actor_1_age, data = subset(dat, decade == d))
  
  # Extract coefficients and add a column for the decade
  coefficients <- data.frame(
    decade = d,
    summary(model)$coefficients[1, 1],
    summary(model)$coefficients[2, 1]
  )
  
  # Combine the results into the tibble
  results_tibble <- bind_rows(results_tibble, as_tibble(coefficients))
}

# Rename the columns for better clarity
colnames(results_tibble) <- c("decade", "Intercept", "slope")

# Print the resulting tibble
results_tibble = results_tibble %>% 
  arrange(desc(slope))
```

The linear model estimates reveal distinctive slopes representing the rate of change in age differences per unit increase in the older actor's age across different decades. Notably, the 1940s, 1970s, and 1950s exhibit steeper slopes (1.13, 1.04, and 0.871, respectively), indicating a more pronounced increase in age differences with advancing older actor ages. In contrast, the 1960s and 2020s present comparatively shallower slopes (0.438 and 0.412, respectively), suggesting a potential attenuation in the rate of increase in age differences during these decades.

# Additional Exploration topic

-   Which directors have the highest average age differences in their movies?

```{r}
dat %>%
  group_by(director) %>%
  summarize(mean_age_diff = mean(age_difference)) %>%
  arrange(desc(mean_age_diff)) %>% 
  head(5) %>% 
  pull(director) -> top_directors
```

Five directors with the highest age difference couples in their movies are `r top_directors`. However, recognizing that many directors have only one couple in their movies, we can refine our analysis by sub-setting the data to include only directors with at least three couples. Afterward, we can calculate the average age difference for this subset of directors.

```{r}
dat %>%
  filter(couple_number>=3) %>% 
  group_by(director) %>%
  summarize(mean_age_diff = mean(age_difference)) %>%
  arrange(desc(mean_age_diff)) %>% 
  head (5) %>% 
  pull(director) -> top_directors3
```

Top 5 directors with the highest age difference and at least 3 couples are `r top_directors3`.

-   We might be curious to see what are the movies? Do they have specific theme?

```{r}
dat %>%
  filter(couple_number>=3) %>% 
  group_by(movie_name) %>%
  summarize(mean_age_diff = mean(age_difference)) %>%
  arrange(desc(mean_age_diff)) %>% 
  head (5) %>% 
  pull(movie_name) -> movie_name
```

Based on the fact that 3 out of 5 movies with the most couples (`r movie_name`) are about the most famous MI6 agent with the licence to kill, we can conclude that James Bond is the Casanova of 20^th^ century.

-   Who is the most active director?

```{r}
directors <- table(dat$director)

top_10 <- head(sort(directors, decreasing = TRUE), 10)
top_10=as.data.frame(top_10)
```

The most active directors are `r top_10$Var1`.

# Conclusion

In conclusion, our analysis of age disparities in cinematic romantic relationships provides valuable insights into the detailed portrayal of such dynamics.

Exploring relationship dynamics and conducting gender-based analyses uncovered significant differences in age gaps between male and female characters. Regression analysis across decades revealed varying trends, with steeper slopes in the 1940s, 1970s, and 1950s, indicating heightened age differences with older actors.

Further investigations into directors with the highest age differences and those consistently portraying such relationships offered additional context. Notably, James Bond movies emerged as movies with the highest number of portrayed romantic relationships.

This project contributes a new perspective to the understanding of age dynamics in cinematic love stories, emphasizing key contextual, gender-related, and temporal influences on on-screen relationships.


# References

1.  Wickham H, Averick M, Bryan J, Chang W, McGowan L, François R, et al. Welcome to the tidyverse. J Open Source Softw. 2019 Nov 21;4(43):1686.
2.  Appelhans T, Detsch F, Nauss T. remote: Empirical Orthogonal Teleconnections in R \[Internet\]. Vol. 65, Journal of Statistical Software. 2015. p. 1--19. Available from: <https://www.jstatsoft.org/article/view/v065i10>
3.  Sievert C. Interactive Web-Based Data Visualization with R, plotly, and shiny \[Internet\]. Chapman and Hall/CRC; 2020. Available from: <https://plotly-r.com>
4.  Csardi G, Nepusz T. The igraph software package for complex network research \[Internet\]. Vol. Complex Systems, InterJournal. 2006. p. 1695. Available from: <https://igraph.org>
5.  Garnier S, Ross N, Rudis R, Camargo PA, Sciaini M, Scherer C. *viridis(Lite) - Colorblind-Friendly Color Maps for R.* \[Internet\]. 2023. Available from: <https://sjmgarnier.github.io/viridis/>. doi:10.5281/zenodo.4679423, viridis package version 0.6.4.
6.  Garnier S, Ross N, Rudis R, Camargo PA, Sciaini M, Scherer C. *viridis(Lite) - Colorblind-Friendly Color Maps for R.* \[Internet\]. 2023. Available from: <https://sjmgarnier.github.io/viridis/>. doi:10.5281/zenodo.4678327, viridisLite package version 0.4.2.
7.  Waring E, Quinn M, McNamara A, Arino de la Rubia E, Zhu H, Ellis S. *skimr: Compact and Flexible Summaries of Data.* 2023. R package version 2.1.5. <https://github.com/ropensci/skimr/>. Available from: <https://docs.ropensci.org/skimr/> (website).
