---
title: "Hollywood age gap"
author: "Mahsa Mohammadi, Luka Milosevic"
format: 
  revealjs:
    code-overflow: wrap
    chalkboard: true
    theme: [solarized]
    code-tools: true
    transition: fade
editor: visual
---

## Data overview

-   Hollywood age gap data - differences in age between movie characters sharing a romantic relationship

```{r}
dat = read.csv("age_gaps.csv")
colnames(dat)
```

## Age distribution 

```{r}
library(ggplot2)
age <- c(dat$actor_1_age, dat$actor_2_age)
age <- as.data.frame(age)

ggplot(age, aes(x=age)) +
  geom_histogram(binwidth=1, position='identity', fill='skyblue', alpha=0.5, color='black') +
  labs(title='Histogram of actors ages', x='Age', y='Frequency') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 
```

## Age gap distribution 
```{r}
library(plotly)


gg <- ggplot(dat, aes(x = age_difference)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Age gap distribution",
       x = "Age difference",
       y = "Frequency") +
  scale_x_continuous(breaks = seq(0, 55, by = 5)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

plotly_plot <- ggplotly(gg)

plotly_plot
```

## Correlation between age and age gap

```{r}
    ggplot(dat, aes(x = actor_1_age, y = age_difference)) +
      geom_point(alpha = 0.7, color = 'skyblue') +
      labs(title = 'Relationship between age and age gap',
           x = 'Actor 1 age',
           y = 'Age Gap') +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5)) 
```

## Trend of the age gap over the years

```{r}
#| echo: false
dat = read.csv("age_gaps.csv")
library(tidyverse)
library(skimr)

dat %>%
  mutate(actor_1_birthdate = ymd(actor_1_birthdate),
         actor_2_birthdate = ymd(actor_2_birthdate), 
         decade = 10 * floor(release_year / 10)) -> dat

dat_long <- dat %>%
  gather(key = "actor", value = "age", actor_1_age, actor_2_age)

ggplot(dat, aes(x = decade, y = age_difference)) +
  geom_point(alpha = 0.7, color = "skyblue", size = 3) +
  geom_smooth(method = "loess", se = TRUE, color = "blue", linetype = "solid", size = 1) +
  labs(title = "Scatter Plot of Decade vs. Age Difference with Smoothed Line",
       x = "Decade",
       y = "Age Difference") +
  scale_x_continuous(breaks = seq(1930, 2020, by = 10)) +
  theme_minimal() +
  theme(panel.grid = element_line(color = "lightgray", linetype = "dashed"))
```

## Network analysis of the specific directors

```{r}
# Filter the DataFrame for a specific director
library(igraph)
director_name <- "Woody Allen"
filtered_data <- dat[dat$director == director_name, ]

# Create a network graph
graph <- graph.data.frame(filtered_data[, c("actor_1_name", "actor_2_name")], directed = FALSE)

# Set up plot parameters
par(mfrow=c(1,1), mar = c(8, 4, 4, 2) + 0.1)  # Adjust the bottom margin (5) as needed
layout <- layout_with_kk(graph)
color_gradient <- viridisLite::viridis(40, option = "plasma")
color_mapping <- color_gradient[cut(filtered_data$age_difference, breaks = 40, labels = FALSE)]

plot(graph, 
     main = paste("Actor Network for Director:", director_name), 
     edge.arrow.size = 0.5,
     vertex.size = 20,
     vertex.label.cex = 0.8,
     layout = layout,
     edge.color = color_mapping
)

# Calculate the legend width based on the graph width
legend_width <- par()$usr[2] - par()$usr[1]  # Graph width

num_colors <- length(color_gradient)
num_columns <- 5  # Adjust the number of columns as needed

legend("bottom", 
       legend = levels(cut(filtered_data$age_difference, breaks = 40)), 
       fill = color_gradient,
       title = "Age Difference",
       cex = 0.7,
       ncol = num_columns,  # Set the number of columns
       xpd = TRUE,  # Allows plotting outside the plot region
       xjust = 0.5,  # Center the legend
       y.intersp = 1.5,  # Adjust the vertical spacing between legend entries
       x = 0,  # Set the x position to place it below the plot
       y = -1.5,  # Set the y position to move it below the plot
       bty = "n")  # Remove the box around the legend

# Reset the xpd parameter to its default value
par(xpd = FALSE)

```

## Network analysis of the specific directors

```{r}
# Filter the DataFrame for a specific director
director_name <- "Guy Hamilton"
filtered_data <- dat[dat$director == director_name, ]

# Create a network graph
graph <- graph.data.frame(filtered_data[, c("actor_1_name", "actor_2_name")], directed = FALSE)

par(mfrow=c(1,1), mar = c(8, 4, 4, 2) + 0.1)  # Adjust the bottom margin (5) as needed

#Choose a layout algorithm

layout <- layout_with_kk(graph)
# Create a color gradient based on age difference
color_gradient <- viridisLite::viridis(40, option = "plasma")

# Map age difference to colors
color_mapping <- color_gradient[cut(filtered_data$age_difference, breaks = 40, labels = FALSE)]

# Plot the network graph with adjusted parameters, layout, and color mapping
plot(graph, 
     main = paste("Actor Network for Director:", director_name), 
     edge.arrow.size = 0.5,
     vertex.size = 20,
     vertex.label.cex = 0.8,
     layout = layout,
     edge.color = color_mapping
)
# Add legend
num_colors <- length(color_gradient)
num_columns <- 5  # Adjust the number of columns as needed

# Add legend with multiple columns
legend("bottom", 
       legend = levels(cut(filtered_data$age_difference, breaks = 40)), 
       fill = color_gradient,
       title = "Age Difference",
       cex = 0.7,
       ncol = num_columns,  # Set the number of columns
       xpd = TRUE,  # Allows plotting outside the plot region
       xjust = 0.5,  # Center the legend
       y.intersp = 1.5,  # Adjust the vertical spacing between legend entries
       x = 0,  # Set the x position to place it below the plot
       y = -1.5,  # Set the y position to move it below the plot
       bty = "n")  # Remove the box around the legend

# Reset the xpd parameter to its default value
par(xpd = FALSE)
```

## Gender analysis

```{r}
gender_age_diff = dat %>% 
       group_by(character_1_gender) %>%
       summarise(mean_age_diff_gender = mean(age_difference))
    man_age_diff = gender_age_diff %>% 
      filter(character_1_gender == "man") 
    woman_age_diff = gender_age_diff %>% 
      filter(character_1_gender =="woman")

    t_test_result = t.test(age_difference ~ character_1_gender, data = dat)
    p.value=t_test_result$p.value
   
    result_object <- list(
  statistics = c(
    t_value = t_test_result$statistic,
    degrees_of_freedom = t_test_result$parameter,
    p_value = t_test_result$p.value
  ),
  means = c(
    mean_group_man = t_test_result$estimate[1],
    mean_group_woman = t_test_result$estimate[2]
  )
)

# Print or use the result_object
print(result_object)
```

## Regression analysis

```{r}
model <- lm(age_difference ~ actor_1_age + decade, data = dat)

ggplot(dat, aes(x = actor_1_age, y = age_difference)) +
  geom_point(color="skyblue") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title = "Prediction of the Age difference based on Older Actor Age through decades",
       x = "Older actor Age",
       y = "Age Difference") +
  facet_wrap(~decade) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Extra investigation about directors
-   Directors with the most number of movies 
```{r}
directors <- table(dat$director)

top_10 <- head(sort(directors, decreasing = TRUE), 10)
top_10=as.data.frame(top_10)
top_10
```
-   Directors with the number of couples at least 3 and and highest age difference

```{r}
dat %>%
  filter(couple_number>=3) %>% 
  group_by(director) %>%
  summarize(mean_age_diff = mean(age_difference)) %>%
  arrange(desc(mean_age_diff)) %>% 
  head (5) %>% 
  pull(director) -> top_directors3
  top_directors3
```

## Movies with the highest number of couples 

```{r}

dat %>%
  filter(couple_number>=3) %>% 
  group_by(movie_name) %>%
  summarize(mean_age_diff = mean(age_difference)) %>%
  arrange(desc(mean_age_diff)) %>% 
  head (5) %>% 
  pull(movie_name) -> movie_name
  movie_name
```

-   James Bond movies 

## Conclusion

- Significant difference between male and female actors in age difference (when one gender is older)

- Age difference tends to increase more with age in the 1940s 

- James Bond movies feature the highest number of couples (3 out of top 5)
